### 2022-06-02 paper/paper.tex
rd: add research and discussion

% Table of Content
\setcounter{tocdepth}{2}

@@

% Links
\usepackage{hyperref}
\hypersetup{
	pdftitle={Development of a finger-key identification submodule for a touch typing trainer},
	pdfstartview={FitH},
	colorlinks=true,
	linkcolor=black,
	citecolor=black,
	filecolor=black,
	urlcolor=MidnightBlue
}

@@

@@

\usepackage{caption}

@@

	\textsc{Development of a finger-key identification submodule\\for a touch typing trainer}
	\par\vspace{0.5cm}

@@

	February 2022
	\vfill
	\hspace{0pt}
\end{titlepage}

@@

\printacronyms{}
\newpage

\listoftables
\newpage

\listoffigures
\newpage

@@

Thus, there is a need for automatically detecting which finger is used during
typing. One way to do this is through finger and hand tracking. One solution for
tracking is by using image processing and machine learning. An example of this
is MediaPipe by~\cite{mediapipe}.

MediaPipe allows for various applications for machine learning in the field of
image processing. This includes, hand tracking, pose estimation, object
detection, and others. Another example of a library that allows for hand and
finger tracking is OpenCV by~\cite{opencv}. This is a tool that simplifies
computer vision and image processing. Machine learning can also be used with
OpenCV.\@

\section{Research Questions}
\begin{enumerate}
	\item What algorithm or technique using computer vision is capable of
	      identifying which keys are being pressed by which fingers during keyboard
	      typing?
	\item What setup and configuration of a single optical camera can be used to
	      capture images during keyboard typing for real-time finger-key identification
	      using the algorithm identified in 1?
	\item How to design and develop a touch typing trainer using the algorithm in 1
	      and camera setup and configuration in 2?
	\item How accurate is the touch typing trainer in identifying which finger is
	      used to press which key?
	\item What new keyboard typing metric can be developed in order to consider both
	      the key and finger used to press as parameters?
\end{enumerate}

@@

\subsection{General Objectives}
To create a touch typing trainer that detects poor finger placement and hand
position to help develop better typing habits and healthier typing ergonomics.

\subsection{Specific Objectives}
\begin{enumerate}
	\item Develop an algorithm or technique using computer vision that is capable
	      of identifying which keys are being pressed by which fingers during keyboard
	      typing --- i.e.\ finger-key identification.
	\item Identify the setup and configuration of a single optical camera that can
	      be used to capture images during keyboard typing for real-time finger-key
	      identification using the algorithm identified in 1
	\item Design and develop a touch typing trainer using the algorithm in 1 and
	      camera setup and configuration in 2
	\item Determine the accuracy of the touch typing trainer in finger-key
	      identification
	\item Develop a new keyboard typing metric that uses both the key and
	      finger used to press as parameters
\end{enumerate}

@@

The setup that will be captured will be simplified and fixed for the development
and testing of the finger-key identification submodule. This includes the
keyboard, surface it is placed on, and lighting of the entire area. Furthermore,
the setup will not be tested by other respondents, but by the researchers only
--- as the focus is on the development and testing of the finger-key
identification submodule and not on the touch typing trainer and its effects on
typing.

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=0.7\textwidth]{60.jpg}
	\caption{A 60\% keyboard in \ac{ansi} layout. Reprinted from \fullcite{60}}
	\label{fig:60}
	\centering
\end{figure}

\section{Significance of the Research}
This research is beneficial for all users of physical keyboards. These include a
vast majority of the population as there are a lot of professions that heavily
rely on keyboards. Examples include developers, physicians, educators, and
accountants. By having better ergonomics while typing --- by touch typing
correctly --- wrist injuries can be prevented, and typing speed may be increased

This research also helps educators, especially early educators teaching beginner
typists. By automatically checking for ergonomics, posture, and correct
technique, the burden of checking each student is lessened, and directed
interventions for bad habits can be easily created as students with these bad
habits are easily identified

This research has a direct impact on people that has hand or wrist injuries that
are caused by poor typing habits. By correcting these poor habits, pain from
these injuries will be lessened, and even be prevented from occurring in the
first place. A specific example of this is by reducing ulnar deviation which
affects the nerve that is indicative of \ac{cts} \parencite{toosi2015}.

@@

% Insert more about keyboard typing here %

@@

\begin{equation}
	WPM = \frac{|T| - 1}{S} \cdot 60 \cdot \frac{1}{5}
\end{equation}

@@

\begin{equation}
	ACC = \frac{|C|}{|T|} \cdot 100\%
\end{equation}

@@

@@

@@

		\node[draw,
			trapezium,
			trapezium left angle = 65,
			trapezium right angle = 115,
			trapezium stretches,
			below=of contours,
			minimum width=3.5cm,
			minimum height=1cm,
			inner sep = 3mm
		] (output) {Key-Edge Coordinatess Map};

@@

\paragraph{Scale Contours}
During testing using training data, a lot of the failures in finger-key
identification was due to the tight fit of the contour to the key. The contour
on its own did not give enough buffer for the placement of the finger. This
buffer was needed as, in some instances, the finger used to press the key was
not exactly on top of the key, but rather to its side. The algorithm accounts
for this situation by adding a buffer. This was achieved through scaling the
contour points by seven pixels on each side. This value was obtained after
testing other possible values. A buffer of five pixels did not have a lot of
effect in reducing the number of failures. A buffer of ten pixels did reduce the
number of failures, but it also greatly increased the algorithm's uncertainty in
determining the finger --- uncertainty is defined as detecting two or more
fingers within one \ac{roi}. Seven was a good middle ground in decreasing
failures, without greatly increasing uncertainty. More information about this
can be found in the Results and Discussion at Section~\ref{section:results}

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=0.8\textwidth]{detected-finger.png}
	\caption{Finger tip within an \ac{roi}}
	\centering
\end{figure}

@@

As such, the expected output of the submodule is the hand and the finger used to
press the key. This allows for a more flexible utilization of the data as it
carries greater context, compared to just returning if the finger used was
correct or not.

@@

\subsection{Implementation}
\label{section:metho-algo-implementation}
The previously discussed algorithms was implemented using Python. The Keyboard
Detection, and Mapping Algorithm in Section~\ref{section:metho-algo-keyboard}
was implemented as a separate submodule in Python using OpenCV. The Finger
Detection, and Tracking Solution in Section~\ref{section:metho-algo-finger} was
consumed using the prebuilt Python package offered by the MediaPipe team.
Finally, the integration of the two will be done as another separate submodule
in Python.

@@

@@

\subsection{Training}
A script was created to test the accuracy of the submodule. This script opened
each video, and the associated labeled file. The script then parsed the labeled
file, and opened the frame for each keypress. The finger-key identification
submodule was then run, using the data needed for the submodule: the key
pressed. The result of the submodule was then compared with the manual
annotation of the keypress. The resulting data from the test was then stored in
a \texttt{.csv} file with the following content:

\begin{table}[H]
	\centering
	\begin{tabular}{ p{0.3\textwidth} p{0.6\textwidth} }
		\hline
		Column Name                           & Definition                                             \\
		\hline
		\texttt{file\_name}                   & Name of the file                                       \\
		\texttt{frame}                        & Frame number                                           \\
		\texttt{key}                          & Key pressed                                            \\
		\texttt{finger}                       & Manual annotation of the finger used to press the key  \\
		\texttt{is\_correct}                  & If the submodule was correct                           \\
		\texttt{fingertips\_detected}         & The fingertips detected by the submodule               \\
		\texttt{keyboard\_detection\_time}    & Time spent in creating the Key-Edge Coordinates Map    \\
		\texttt{finger\_classification\_time} & Time spent in classifying which finger pressed the key \\
	\end{tabular}
	\caption{\label{tab:csv-output}CSV File Contents}
\end{table}

Adjustments were then made to the algorithm after each training pass to improve
its performance. Some runs had improvements in its statistics without doing any
adjustments. These were runs that showed errors in the manual annotation and
were corrected afterwards.

@@

\begin{table}[H]
	\centering
	\begin{tabular}{ p{0.3\textwidth} p{0.6\textwidth} }
		\hline
		Identifications                                                                                                                       \\
		\hline
		Success                                     & Number of successful finger-key identifications                                         \\
		Failed                                      & Number of failed finger-key identifications                                             \\
		Success Percentage                          & Percentage of successful finger-key identifications                                     \\
		Failed Percentage                           & Percentage of failed finger-key identifications                                         \\
		\\[0.25cm]
		\hline
		Failure Types                                                                                                                         \\
		\hline
		Mismatched Identification                   & Submodule output did not match manual annotation                                        \\
		No Identification                           & Submodule did not detect finger tips in \ac{roi}                                        \\
		\\[0.25cm]
		\hline
		Uncertain Successes                                                                                                                   \\
		\hline
		Total                                       & Total uncertain successes, defined as detecting two or more fingers within one \ac{roi} \\
		Spacebar                                    & All uncertain successes that were on the spacebar                                       \\
		Spacebar Percentage                         & Percentage of spacebar uncertain success over all uncertain successes                   \\
		Non-spacebar                                & All other uncertain successes                                                           \\
		Non-spacebar Percentage                     & Percentage of non-spacebar uncertain success over all uncertain successes               \\
		Non-spacebar over total \newline Percentage & Percentage of non-spacebar uncertain success over all total keypresses                  \\
		\\[0.25cm]
		\hline
		Average Running Time                                                                                                                  \\
		\hline
		Keyboard Detection Time                     & Average keyboard detection time of all videos                                           \\
		Finger Identification Time                  & Average finger identification time of all key presses                                   \\
	\end{tabular}
	\caption{\label{tab:acquired-stats}Acquired statistics}
\end{table}

@@

%% \subsubsection{Finger-key Identification}
%% The submodule that integrates all the algorithms will be called within the
%% framework. The framework will then respond to requests for finger-key
%% identification using this submodule. This will allow the back-end to expose its
%% finger-key identification abilities and allow the front-end to use this data for
%% the trainer.

@@

\section{Metric}
\label{section:metric}

@@

\subsection{Per Test Sequence}
The metric which calculates per test sequence is \ac{fpacc}. This metric
computes the percentage of accurate keys pressed with the correct finger over the
length of the typing test sequence. Inputting the wrong character, even if
pressed with the correct finger, reduces \ac{fpacc} since \ac{fpacc} measures
\emph{accurate key presses}, and incorrect characters areconsidered as
inaccurate key presses.

@@

\begin{equation}
	FP ACC = \frac{|F|}{|T|} \cdot 100\%
\end{equation}

@@

\subsection{Per Key}
This metric computes the \ac{hfpacc} for a certain key. This metric takes the
times the user pressed the wrong key when asked for a certain key into account,
the same as in \ac{fpacc}.\@This allows the user to easily verify which keys
need attention and retraining if there is a high frequency of error.

@@

\begin{equation}
	HFP ACC = \frac{|F_{char}|}{|C_{char}|} \cdot 100\%
\end{equation}

@@

\chapter{Results and Discussion}
\chapter{Conclusion}
\chapter{Future Work}

>>>>

@@

% Links
\usepackage[linktoc=all]{hyperref}
\hypersetup{
	pdftitle={Development of a Finger-Key Identification Module for a Touch Typing Trainer},
	pdfstartview={FitH},
	colorlinks=true,
	linkcolor=black,
	citecolor=black,
	filecolor=black,
	urlcolor=MidnightBlue
}

@@

% Graphs
\usepackage{pgfplots}
\usepgfplotslibrary{external}
\pgfplotsset{width=1\textwidth,compat=1.9}

@@

% Center table with controlled width
\usepackage{array}
\newcolumntype{P}[1]{>{\centering\arraybackslash}p{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

% List of Equations
\usepackage[titles]{tocloft}
\newcommand{\listequationsname}{List of Equations}
\newlistof{equations}{equ}{\listequationsname}
\newcommand{\labelequations}[1]{%
	\addcontentsline{equ}{equations}{\numberline{\theequation}#1}\par}
\setlength{\cftequationsindent}{1.5em}
\setlength{\cftequationsnumwidth}{2.3em}

\addto\captionsamerican{% Replace "english" with the language you use
	\renewcommand{\contentsname}%
	{Table of Contents}%
}

@@

	\textsc{Development of a Finger-Key Identification Module\\for a Touch Typing Trainer}
	\par\vspace{0.5cm}

@@

	June 2022
	\vfill
	\hspace{0pt}
\end{titlepage}

@@

\begin{center}

	\includegraphics[width=0.2\textwidth]{upc.png}
	\includegraphics[width=0.2\textwidth]{dcs.png}
	\par\vspace{0.25cm}

	\textbf{\uppercase{University of the Philippines Cebu}}\\
	College of Science\\
	Department of Computer Science

	\par\vspace{1cm}

	\textbf{\uppercase{Development of a finger-key identification module\\for a touch typing trainer}}

	\par\vspace{1cm}

	\textbf{Permission is given for the following people to have access to this thesis:}

	\par\vspace{0.25cm}

	\begin{tabular}{ | P{0.8\textwidth} | P{0.1\textwidth} | }
		\hline
		Available to the general public                                 & Yes \\
		\hline
		Available only after consultation with author or thesis adviser & Yes \\
		\hline
		Available to those bound by confidentiality agreement           & Yes \\
		\hline
	\end{tabular}
	\par\vspace{1cm}

	\line(1,0){0.6\textwidth}\\
	Oscar Vian L. Valles\\
	Student
	\par\vspace{0.5cm}

	\line(1,0){0.6\textwidth}\\
	Dhong Fhel K. Gom-os\\
	Special Problem Adviser

	\vfill
	\hspace{0pt}

\end{center}

\noindent
This thesis entitled ``\textbf{\uppercase{Development of a finger-key
		identification \\module for a touch typing trainer}}'' prepared and
submitted by \\\uppercase{Oscar Vian L. Valles} is approved by:

\par\vspace{2cm}

\begin{center}
	\begin{minipage}[t]{0.45\textwidth}
		\centering
		\line(1,0){0.9\textwidth}\\
		Prof. Aileen Joan O. Vicente\\
		Panel Member

		\par\vspace{0.5cm}
		\line(1,0){0.9\textwidth}\\
		Date
	\end{minipage}
	\begin{minipage}[t]{0.45\textwidth}
		\centering
		\line(1,0){0.9\textwidth}\\
		Prof. Aileen Joan O. Vicente\\
		Panel Member

		\par\vspace{0.5cm}
		\line(1,0){0.9\textwidth}\\
		Date
	\end{minipage}

	\par\vspace{2cm}

	\line(1,0){0.405\textwidth}\\
	Prof. Aileen Joan O. Vicente\\
	Panel Chair

	\par\vspace{0.5cm}
	\line(1,0){0.405\textwidth}\\
	Date

	\par\vspace{2cm}

	\begin{minipage}{0.45\textwidth}
		\centering
		\line(1,0){0.9\textwidth}\\
		Prof. Aileen Joan O. Vicente\\
		Chair, Department of Computer Science

		\par\vspace{0.5cm}
		\line(1,0){0.9\textwidth}\\
		Date
	\end{minipage}
	\begin{minipage}{0.45\textwidth}
		\centering
		\line(1,0){0.9\textwidth}\\
		Prof. Nelia S. Ereno\\
		Dean, College of Science

		\par\vspace{0.5cm}
		\line(1,0){0.9\textwidth}\\
		Date
	\end{minipage}
\end{center}

\pagenumbering{roman}
\addcontentsline{toc}{chapter}{Abstract}
\begin{center}
	\LARGE\textbf{Abstract}
\end{center}

\addcontentsline{toc}{chapter}{Acronyms}
\printacronyms{}
\newpage

\addcontentsline{toc}{chapter}{List of Tables}
\listoftables
\newpage

\addcontentsline{toc}{chapter}{List of Figures}
\listoffigures
\newpage

\addcontentsline{toc}{chapter}{List of Equations}
\listofequations

@@

Thus, there is a need for automatically identifying which finger is used to
press which key during typing --- hereby defined as finger-key identification.

\section{Research Questions}
\begin{enumerate}
	\item What algorithm or technique using computer vision is capable of
	      finger-key identification.
	\item What setup and configuration of a single optical camera can be used to
	      capture images during keyboard typing for time finger-key identification
	      using the algorithm identified in 1?
	\item How to design and develop a finger-key identification module using
	      the algorithm in 1 and camera setup and configuration in 2?
	\item How accurate is the module developed in 3 in finger-key identification?
	\item Is the module developed in 3 fast enough to run in real-time?
	\item What new keyboard typing metric to develop that considers both the key and
	      finger used to press as parameters?
\end{enumerate}

@@

\subsection{General Objectives}
To create a finger-key identification module that accurately identifies what
finger was used to press a key at a point in time to help develop better typing
habits and healthier typing ergonomics.

\subsection{Specific Objectives}
\begin{enumerate}
	\item Develop an algorithm or technique using computer vision that is capable
	      of finger-key identification.
	\item Identify the setup and configuration of a single optical camera that can
	      be used to capture images during keyboard typing for finger-key
	      identification using the algorithm identified in 1
	\item Design and develop a module using the algorithm in 1 and camera setup
	      and configuration in 2
	\item Determine the accuracy of the module developed in 3 in finger-key
	      identification
	\item Determine if the module developed in 3 is fast-enough to run in
	      real-time
	\item Develop a new keyboard typing metric that uses both the key and
	      finger used to press as parameters
\end{enumerate}

@@

The setup that will be captured will be simplified and fixed for the development
and testing of the finger-key identification module. This includes the
keyboard, surface it is placed on, and lighting of the entire area. Furthermore,
the setup will not be tested by other respondents, but by the researchers only
--- as the focus is on the development and testing of the finger-key
identification module and not on the touch typing trainer and its effects on
typing.

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=0.7\textwidth]{60.png}
	\caption{A 60\% keyboard in \ac{ansi} layout.}
	\label{fig:60}
	\centering
\end{figure}

\section{Significance of the Research}
This research is beneficial for all users of physical keyboards. These include a
majority of the population as there are a lot of professions that heavily rely
on keyboards. Examples include developers, physicians, educators, and
accountants. By having better ergonomics while typing --- by touch typing
correctly --- wrist injuries can be prevented, and typing speed may be increased

This research also helps educators, especially early educators teaching beginner
typists. By automatically checking for correct technique, the burden of checking
each student is reduced, and directed interventions for bad habits can be
easily created as students with these bad habits are easily identified

This research has a direct impact on people that have hand or wrist injuries
that are caused by poor typing habits. By correcting these poor habits, pain
from these injuries will be reduced, and even be prevented from occurring in the
first place. A specific example of this is by reducing ulnar deviation which
affects the nerve that is indicative of \ac{cts} \parencite{toosi2015}.

@@

@@

\begin{equation}
	WPM = \frac{|T| - 1}{S} \cdot 60 \cdot \frac{1}{5}
	\label{equ:wpm}
\end{equation}
\labelequations{Words per Minute}

@@

\begin{equation}
	ACC = \frac{|C|}{|T|} \cdot 100\%
\end{equation}
\labelequations{Accuracy}

@@

\subsubsection{Average Typing Speed}

Average typing speed using a keyboard with QWERTY layout, based on test data
from \cite{keybr}, is $\approx37.5$ \ac{wpm}. This average is taken from all of
the test results they have gathered from all of their users.
\cite{arif2009} also performed a survey with 12 participants using
multiple typing methods. They got an average \ac{wpm} of 75.85 with a standard
deviation of 15.61 when their participants used a physical QWERTY layout. This
average \ac{wpm} was also the highest among the other typing methods the
researchers studied.

@@

\subsection{Computer}
The computer used to run the algorithm is a desktop computer with the following
specification:

\begin{table}[H]
	\small
	\centering
	\begin{tabular}{ p{0.3\textwidth} p{0.6\textwidth} }
		\toprule
		Component & Specification                          \\
		\midrule
		CPU       & AMD Ryzen 5 3600                       \\
		GPU       & AMD Radeon RX 5600 XT                  \\
		RAM       & G.Skill Trident Z Neo RGB 16GB 3200mhz \\
		Storage   & Samsung SSD 850 EVO                    \\
		OS        & Arch Linux                             \\
		\bottomrule
	\end{tabular}
	\caption{\label{tab:table-specs}Computer Specifications}
\end{table}

@@

		\node[draw,
			trapezium,
			trapezium left angle = 65,
			trapezium right angle = 115,
			trapezium stretches,
			below=of contours,
			minimum width=3.5cm,
			minimum height=1cm,
			inner sep = 3mm
		] (output) {Key-Edge Coordinates Map};

@@

\paragraph{Scale Contours}
During testing using training data, a lot of the failures in finger-key
identification was due to the tight fit of the contour to the key. The contour
on its own did not give enough buffer for the placement of the finger. This
buffer was needed as, in some instances, the finger used to press the key was
not exactly on top of the key, but rather to its side. The algorithm accounts
for this situation by adding a buffer. This was achieved through scaling the
contour points by seven pixels on each side. This value was obtained after
testing other possible values. A buffer of five pixels did not have a lot of
effect in reducing the number of failures. A buffer of ten pixels did reduce the
number of failures, but it also greatly increased the algorithm's uncertainty in
determining the finger --- uncertainty is defined as detecting two or more
fingers within one \ac{roi}. Seven was a good middle ground in decreasing
failures, without greatly increasing uncertainty.

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=0.8\textwidth]{detected-finger.png}
	\caption{Fingertip within an \ac{roi}}
	\centering
\end{figure}

@@

As such, the expected output of the module is the hand and the finger used to
press the key. This allows for a more flexible utilization of the data as it
carries greater context, compared to just returning if the finger used was
correct or not.

@@

\subsection{Implementation}
\label{section:metho-algo-implementation}
The previously discussed algorithms was implemented using Python. The Keyboard
Detection, and Mapping Algorithm in Section~\ref{section:metho-algo-keyboard}
was implemented as a separate module in Python using OpenCV. The Finger
Detection, and Tracking Solution in Section~\ref{section:metho-algo-finger} was
consumed using the prebuilt Python package offered by the MediaPipe team.
Finally, the integration of the two will be done as another separate module
in Python.

@@

One caveat of the camera used was it's autofocus capabilities. This meant that
there was a set amount of time during the start of the recording where the
camera captured blurry images of the keyboard. This necessitates around three
seconds for the autofocus to complete and focus on the keyboard. There was one
video capture where the whole test sequence was scrapped as the researcher
started typing even if the camera was not focused.

@@

\subsection{Training}
A script was created to test the accuracy of the module. This script opened each
video, and the associated labeled file. There were three frames taken for each
video to perform step one of the module: frames ten, fifteen, and twenty. The
first frame that returned a Key-Edge Coordinates Map after going through step
one of the module was used. The rest of the frames were not passed through the
module and were discarded.

The script then parsed the labeled file, and opened the frame for each keypress.
The second algorithm of the module was then run, using the data needed for the
algorithm: the key pressed. The result of the module was then compared with the
manual labelling of the keypress. The resulting data from the test was then
stored in a \texttt{.csv} file with the following content:

\begin{table}[H]
	\small
	\centering
	\begin{tabular}{ p{0.3\textwidth} p{0.6\textwidth} }
		\toprule
		Column Name                           & Definition                                             \\
		\midrule
		\texttt{file\_name}                   & Name of the file                                       \\
		\texttt{frame}                        & Frame number                                           \\
		\texttt{key}                          & Key pressed                                            \\
		\texttt{finger}                       & Manual labelling of the finger used to press the key   \\
		\texttt{is\_correct}                  & If the module was correct                              \\
		\texttt{fingertips\_detected}         & The fingertips detected by the module                  \\
		\texttt{keyboard\_detection\_time}    & Time spent in creating the Key-Edge Coordinates Map    \\
		\texttt{finger\_classification\_time} & Time spent in classifying which finger pressed the key \\
		\bottomrule
	\end{tabular}
	\caption{\label{tab:csv-output}CSV File Contents}
\end{table}

Adjustments were then made to the algorithm after each training pass to improve
its performance. Some runs had improvements in its statistics without doing any
adjustments. These were runs that showed errors in the manual labelling and
were corrected afterward.

@@

\begin{table}[H]
	\small
	\centering
	\begin{tabular}{ p{0.3\textwidth} p{0.6\textwidth} }
		\toprule
		Identifications                                                                                                                    \\
		\midrule
		Successes                                & Number of successful finger-key identifications                                         \\
		Failures                                 & Number of failed finger-key identifications                                             \\
		Success Percentage                       & Percentage of successful finger-key identifications                                     \\
		Failure Percentage                       & Percentage of failed finger-key identifications
		\\[0.25cm]
		\midrule
		Failure Types                                                                                                                      \\
		\midrule
		Mismatched Identification                & Module output did not match manual labelling                                            \\
		No Identification                        & Module did not detect fingertips in \ac{roi}
		\\[0.25cm]
		\midrule
		Uncertain Successes                                                                                                                \\
		\midrule
		Total                                    & Total uncertain successes, defined as detecting two or more fingers within one \ac{roi} \\
		Spacebar                                 & All uncertain successes that were on the spacebar                                       \\
		Spacebar Percentage                      & Percentage of spacebar uncertain successes over all uncertain successes                 \\
		Spacebar Percentage (Total)              & Percentage of spacebar uncertain successes over total keypresses                        \\
		Non-spacebar                             & All other uncertain successes                                                           \\
		Non-spacebar Percentage                  & Percentage of non-spacebar uncertain successes over all uncertain successes             \\
		Non-spacebar Percentage \newline (Total) & Percentage of non-spacebar uncertain successes over total keypresses
		\\[0.25cm]
		\midrule
		Average Running Time                                                                                                               \\
		\midrule
		Keyboard Detection Time                  & Average keyboard detection time of all videos                                           \\
		Finger Identification Time               & Average finger identification time of all key presses                                   \\
		\bottomrule
	\end{tabular}
	\caption{\label{tab:acquired-stats}Acquired statistics}
\end{table}

@@

%% \subsubsection{Finger-key Identification}
%% The module that integrates all the algorithms will be called within the
%% framework. The framework will then respond to requests for finger-key
%% identification using this module. This will allow the back-end to expose its
%% finger-key identification abilities and allow the front-end to use this data for
%% the trainer.

@@

\section{Metrics}
\label{section:metric}

@@

\subsection{Per Test Sequence}
The metric which calculates per test sequence is \ac{fpacc}. This metric
computes the percentage of accurate keys pressed with the correct finger over the
length of the typing test sequence. Inputting the wrong character, even if
pressed with the correct finger, reduces \ac{fpacc} since \ac{fpacc} measures
\emph{accurate key presses}, and incorrect characters are considered as
inaccurate key presses.

@@

\begin{equation}
	FP ACC = \frac{|F|}{|T|} \cdot 100\%
\end{equation}
\labelequations{Finger Placement Accuracy}

@@

\subsection{Per Key}
This metric computes the \ac{hfpacc} for a certain key. This metric takes the
times the user pressed the wrong key when asked for a certain key into account,
the same as in \ac{fpacc}. This allows the user to easily verify which keys
need attention and retraining if there is a high frequency of error.

@@

\begin{equation}
	HFP ACC = \frac{|F_{char}|}{|C_{char}|} \cdot 100\%
\end{equation}
\labelequations{Historical Finger Placement Accuracy}

@@

\chapter{Results and Discussion}
\label{section:rd}

\section{Accuracy}
Accuracy, in the context of the finger-key identification module calculates the
percentage of successful finger-key identifications that the module has
performed. In this case, success is defined as having the expected finger, based
on the manual labelling, equal the return value of the module. Accuracy is
calculated as

\begin{equation}
	ACC = \frac{|S|}{|T|}
\end{equation}
\labelequations{Finger-Key Identification Module Accuracy}

Where $|S|$ is the number of successful finger-key identifications, and $|T|$ is
the total number of keypresses for all typing test sequences. This is also
referred to as the success percentage.

\subsection{Training Results}
There were a total of seven training iterations. These training iterations
served as a way to improve the performance of the module by adjusting its
parameters.

\subsubsection{Identifications}

\begin{figure}[H]
	\centering
	\begin{tikzpicture}
		\begin{axis}[
				ybar stacked,
				bar width=30pt,
				xlabel={Training Iteration},
				ylabel={Percentage over total keypresses},
				xmin=1, xmax=7,
				ymin=86, ymax=100,
				xtick={1,2,3,4,5,6,7},
				ytick={86, 86, 88, 90, 92, 94, 96, 98, 100},
				enlargelimits=0.15,
				area legend,
				tick label style={font=\footnotesize},
				legend style={font=\footnotesize},
				label style={font=\footnotesize},
				every axis legend/.append style={cells={anchor=west}},
				ymajorgrids=true,
			]

			\addplot[
				CornflowerBlue,
				fill=CornflowerBlue,
			]
			coordinates {
					(1, 88.47) (2, 88.95) (3, 98.51) (4, 99.55) (5, 98.47) (6, 99.17) (7, 99.49)
				};
			\addplot[
				Salmon,
				fill=Salmon,
			]
			coordinates {
					(1, 11.46) (2, 10.98) (3, 1.42) (4, 0.38) (5, 1.46) (6, 0.76) (7, 0.45)
				};
			\legend{Successes, Failures}

		\end{axis}
	\end{tikzpicture}
	\caption{Identification results of the training iterations}
	\label{fig:rd-training-identification}
\end{figure}

Over the seven training iterations, there was an increase in success percentage.
Consequently, there was also a decrease in failure percentage. There was a total
of 1475 total keypresses for iterations one to three, and 1571 total keypresses
for iteration four to seven.

The increase in total keypresses starting from iteration four was due to adding
multiple tries to the training routine in obtaining the Key-Edge Coordinates
Map. Iterations one through three only tried frame 10 as the source frame for
the module while iterations four through seven tried frame 10, 15, and 20.

There was a need to try different frames since the camera was not consistent in
its ability to focus on the keyboard for each video. In some videos video, the
camera was not focused on the keyboard on frame 10, but it was able to focus on
frame 15. As a result, the total number of available keypresses to test
increased.

The 0.48\% jump between iteration one and two can be attributed to correcting
erroneous finger-key identifications that was obtained from manual labelling.
The 9.56\% jump in success percentage between iteration two and three was due to
adding buffers around the \ac{roi} by scaling the contours by 10 pixels on each
side. Iteration four's increase was due to fixing the image map. There were
certain keys that had incorrect color values assigned in the Key-Color Values
Map, and some keys were of incorrect shape. Iterations five to seven were
performed to test different values and shapes for scaling the \ac{roi}.

\subsubsection{Uncertain Successes}

\begin{figure}[H]
	\centering
	\begin{tikzpicture}
		\begin{axis}[
				ybar stacked,
				bar width=30pt,
				xlabel={Training Iteration},
				ylabel={Percentage over total keypresses},
				xmin=1, xmax=7,
				ymin=0, ymax=30,
				xtick={1,2,3,4,5,6,7},
				ytick={0,5,10,15,20,25,30},
				enlargelimits=0.15,
				area legend,
				tick label style={font=\footnotesize},
				legend style={font=\footnotesize},
				label style={font=\footnotesize},
				every axis legend/.append style={cells={anchor=west}},
				ymajorgrids=true,
			]

			\addplot[
				CornflowerBlue,
				fill=CornflowerBlue
			]
			coordinates {
					(1, 5.69) (2, 5.69) (3, 12.61) (4, 12.86) (5, 9.32) (6, 10.76) (7, 12.86)
				};
			\addplot[
				Salmon,
				fill=Salmon,
			]
			coordinates {
					(1, 0.61) (2, 0.61) (3, 17.15) (4, 17.06) (5, 3.95) (6, 6.49) (7, 9.68)
				};
			\legend{Spacebar, Non-spacebar}

		\end{axis}
	\end{tikzpicture}
	\caption{Uncertain Successes of the training iterations}
	\label{fig:rd-training-uncertain-successes}
\end{figure}

Uncertain Successes are instances where the finger-key identification module was
able to detect two or more fingertips within the \ac{roi} of the key. As such,
the module is uncertain which of the multiple fingertips actually pressed the
key. However, it is still classified as a success since the expected finger from
the manual labelling exists within the array of fingertips returned by the
module.

\paragraph{Spacebar}
In a 60\% keyboard, the spacebar is the longest key on the keyboard. In
addition, this key is right below the non-dominant hand's thumb default resting
position when the hand's posture follows the proper touch typing posture. As
such, it was to be expected that a majority of the uncertain successes were
because of the spacebar. However, this was not the case for iterations three and
four. This was because the buffer for the \ac{roi} was too big and this caused
the module to detect more than one fingertip within that \ac{roi}.

\paragraph{Non-spacebar}

There were cases where the module had uncertain successes that were not over the
spacebar. In almost practically all cases, these were over the alphanumeric
keys. This uncertain successes stem from the increased buffer of the \ac{roi}.
This type of uncertain successes carry more weight compared to spacebar
uncertain successes since the keys where these uncertain successes come from are
the smallest keys in the keyboard. This means that the chances that a person
actually has two or more fingers over the keys is rare, compared to the
spacebar.

\subsubsection{Uncertain Successes vs Success Percentage}
\label{section:rd-uncertain vs success}

\begin{figure}[H]
	\centering
	\begin{tikzpicture}
		\begin{axis}[
				xlabel={Training Iteration},
				ylabel={Percentage over total keypresses},
				xmin=1, xmax=7,
				ymin=0, ymax=100,
				xtick={1,2,3,4,5,6,7},
				ytick={0,20,40,60,80,100},
				tick label style={font=\footnotesize},
				legend style={font=\footnotesize},
				label style={font=\footnotesize},
				enlargelimits=0.15,
				ymajorgrids=true,
				xmajorgrids=true,
				nodes near coords,
				every node near coord/.append style={font=\footnotesize},
				every axis legend/.append style={at={(0.98, 0.02)}, anchor=south east, cells={anchor=west}}
			]

			\addplot[
				color=CornflowerBlue,
				mark=square,
			]
			coordinates {
					(1, 88.47) (2, 88.95) (3, 98.51) (4, 99.55) (5, 98.47) (6, 99.17) (7, 99.49)
				};
			\addplot[
				color=Salmon,
				mark=square,
			]
			coordinates {
					(1, 6.31) (2, 6.31) (3, 29.76) (4, 29.92) (5, 13.18) (6, 17.25) (7, 22.53)
				};

			\legend{Successes, Uncertain Successes}

		\end{axis}
	\end{tikzpicture}
	\caption{Uncertain Successes of the training iterations}
	\label{fig:rd-training-uncertain-successes}
\end{figure}

Tuning the algorithm required managing both uncertain successes and the success
percentage and making compromises between one or the other. The goal of the
training process was to maximize the success percentage, while limiting the
number of uncertain successes.

It can be noted that in iterations one and two, the success rate was lower
compared to the other iterations. However, the percentage of uncertain successes
was also the lowest. Iteration three was the first iteration that created a
buffer around each key, however this buffer of 10 pixels was too aggressive,
with the uncertain successes spiking upwards to nearly 30\%. Iteration five
reduced the buffer size in half to five pixels. This did not affect success rate
that much, with only a reduction of 1.07\%. This also greatly decreased the
uncertain success percentage by 16.74\%. The sixth iteration met halfway between
iteration four and five, by adding a buffer of seven pixels. This increased the
success rate back to 99.17\% without greatly increasing the number of uncertain
successes. In comparison with the fourth iteration, the sixth iteration had its
success rate decrease by 0.38\%, but its percentage of uncertain successes
decreased by 12.67\%. A seventh iteration was trialed where the increase in
buffer was asymmetrical. For each key, the points nearer the user were increased
by 10 pixels, while the points farther away were increased by five pixels only.
This slightly increased the success rate by 0.32\%. However, the percentage of
uncertain successes also increased by 5.28\%.

Based on these results, the final selected parameters for the module was the
parameters set in iteration six, with the buffer set to seven pixels as it was a
good middle ground between success rate, and uncertain successes.

\subsubsection{Failure Types}
